---
title: "Regression Tree"
author: "Francia Moscoso"
date: "May 1, 2016"
output: html_document
---

```{r LoadLib, message=F, warning=F}
library(dplyr)     
library(corrplot) 
library(ggplot2)
library(gridExtra)
library(rpart)
library(rpart.plot)
library(caret)
library(e1071)
```
<br>
<br>
**Loading Data Sets**
```{r comment="", echo=TRUE}
# Working Directory
setwd("~/SprintboardProject/PrudentialIns/RegressionTree") 

train <- read.csv("../DataSets/train.csv", header = TRUE)
test <-  read.csv("../DataSets/test.csv", header = TRUE)
```
<br>
<br>
**Function that populates missing values with the median**    
```{r comment="", echo=TRUE}
manage_na <- function(tempo)
{
  for(i in 1:ncol(tempo))
  {
    if(is.numeric(tempo[,i]))
    {
      tempo[is.na(tempo[,i]),i] <- median(tempo[!is.na(tempo[,i]),i])
    }
  }
  tempo
}
```
<br>
**Analyze dependent variable Train Response** 
```{r comment="", echo=TRUE}
nrow(train) 

summary(train$Response)

table(train$Response)

```
<br>
<br>
**Applying Regression Tree method**
```{r comment="", echo=TRUE}
# We can run 'rpart' with na values without errors. 
RegTree <- rpart(Response ~ ., data=train)
prp(RegTree)
```
<br>
**The Regression tree does not have many leaves and it is easy to interpretate. However, I don't see outputs for Responses closer to 7 or 8**
<br>
<br>
**Making Predictions on test data set**
```{r comment="", echo=TRUE}
RegTree.pred = predict(RegTree, newdata = test)

table(RegTree.pred)
```
**The Tree for the prediction on the Test data set does not have many leaves and it is easy to interpretate. However, I don't see outputs for Responses closer to 7 or 8**
<br>
<br>
**Populate train$Response missing values with the median so we won't get errors in Cross Validation**
```{r comment="", echo=TRUE}
train <- manage_na(train[,-c(1)])     #   Except columns 1 (ID)    

test <- manage_na(test[,-c(1)])

```
<br>
<br>
**Cross Validation**
```{r comment="", echo=TRUE}
tr.control <- trainControl(method="cv", number=10)
cp.grid <- expand.grid(.cp=1:10)*0.001

train$Product_Info_2 = NULL  # OTherwise, it will give error when predicting on test data set

#train will give error if the data has na values
tr <- train(Response ~ ., data=train, method = "rpart", trControl = tr.control, tuneGrid = cp.grid)

tr
```
**The final value used for the model was cp = 0.001 with the best root mean square error(RMSE=1.975)**
<br>
<br>
**Plot the best tree**
```{r comment="", echo=TRUE}
best.tree <- tr$finalModel

# 65535 nodes
best.tree

prp(best.tree)

## test$Product_Info_2 = NULL Gives error categorical variables
best.tree.pred = predict(best.tree, newdata=test)
  
```
<br>
<br>
**Predicting best tree on test data set**
```{r comment="", echo=TRUE}
## test$Product_Info_2 = NULL - Gives error categorical variables
best.tree.pred = predict(best.tree, newdata=test)

#Distribution of Response predicted on Test data set 
summary(best.tree.pred)
```
<br>
<br>